---
title: "National Diet and Nutrition Survey_DiD"
author: "Constanza Avalos"
date: "2024-04-15"
output: 
  html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Datasets 2008-2019

This survey contains various datasets, with the food diary report being the focus of my analysis. The food diary report is a self-reported record of all items consumed over a 24-hour period (midnight to midnight) on the previous day. Participants were required to complete four recalls on different days that were not consecutive. There are three primary datasets: one categorizing consumption reports into food categories, another detailing grams consumed per category at an individual level, and finally one indicating nutrient intake at the individual level. These databases include information such as reporting days of the week, time, date, country, age, gender and others. Moreover, they have a multilevel structure encompassing individuals, households, regions and countries.The data also encompasses socio-demographic details along with anthropometric measurements (BMI, weight, height) at an individual level.

## Treatment groups

The dataset "ndns_rp_yr9_yr9_11a_dayleveldietarydata_nutrients_uk_20210831" contains information on nutrient intake at the individual level. I plan to divide the data for analysis, specifically focusing on individuals aged 18 and older. Additionally, this database includes the variable region, which is used to differentiate between the treatment and control groups.

Individuals in the lower income_tertile (D=1) are included in the treatment group, while others are placed in the control group (D=0) as part of the second quasi-experimental design. Previous research suggests that lower socioeconomic groups have benefited most since retailers implemented this policy. Unlike taxes, food labeling does not place an economic burden on consumers but instead guides them toward healthy eating behavior.

```{r data, echo=FALSE}

# Working directory

setwd("C:/Users/n51867ca/OneDrive - The University of Manchester/R/Second paper")

# input Stata file
library(openxlsx)
library(readxl)


library(readr)
DS_individuals_exploration_10042024_full <- read_csv("DS_individuals_exploration_10042024_full.csv")


# Data structure

head(DS_individuals_exploration_10042024_full)

```

## Statistics 

```{r summary, echo=FALSE, results='asis'}

library(summarytools)

# Replace negative values with NA across the entire dataset
DS_individuals_exploration_10042024_full[DS_individuals_exploration_10042024_full < 0] <- NA

summary <- dfSummary(DS_individuals_exploration_10042024_full, graph.magnif = 0.75, plain.ascii = FALSE, style = "grid")
# Open a PDF device

print(summary, method = 'render')
```

## Identification Treatment Groups - Income Tertiles

```{r DiD, echo=FALSE}

library(tidyverse)
library(lmtest)
library(dplyr)

DS_individuals_exploration_10042024_full <- DS_individuals_exploration_10042024_full %>%
    mutate(Year = as.numeric(Year), # Convert SurveyYear to numeric
          Year = case_when(
           Year == 1 ~ 2009,
           Year == 2 ~ 2010,
           Year == 3 ~ 2011,
           Year == 4 ~ 2012,
           Year == 5 ~ 2013,
           Year == 6 ~ 2014,
           Year == 7 ~ 2015,
           Year == 8 ~ 2016,
           Year == 9 ~ 2017,
           Year == 10 ~ 2018,
           Year == 11 ~ 2019,
    TRUE ~ Year # Keeps the original value if none of the conditions above are met
  ))


# Convert -1 in Body Mass Index (bmival2) to NA, create treatment variable
DS_individuals_exploration_10042024_full <- DS_individuals_exploration_10042024_full %>%
    filter(Income != -1) %>%
    mutate(
    treatment = if_else(Income > 1, 0, 1), # Include condition here with the condition that defines the treatment group
    post_policy = if_else(Year >= 2013, 1, 0)
  )

#create a contingency table
cont_table <- table(DS_individuals_exploration_10042024_full$Year, DS_individuals_exploration_10042024_full$treatment)
cont_table
percent_table <- prop.table(cont_table) * 100
percent_table
cont_table_bmi <- table(DS_individuals_exploration_10042024_full$Year, DS_individuals_exploration_10042024_full$Income)
cont_table_bmi
cont_table_pol <- table(DS_individuals_exploration_10042024_full$Year, DS_individuals_exploration_10042024_full$post_policy)
cont_table_pol

```

## Weighting

```{r weight, echo=FALSE}

# Multiply each weight by 15655

DS_individuals_exploration_10042024_full$adjustedweight <- DS_individuals_exploration_10042024_full$Weight * 15655

#Multiply each by the combined sum of the four weights

DS_individuals_exploration_10042024_full$FinalWeight <- ifelse(DS_individuals_exploration_10042024_full$Year >= 2009 & DS_individuals_exploration_10042024_full$Year <= 2012, 
                           (DS_individuals_exploration_10042024_full$adjustedweight / 6828) * 0.3636,
                           ifelse(DS_individuals_exploration_10042024_full$Year >= 2013 & DS_individuals_exploration_10042024_full$Year <= 2014,
                                  (DS_individuals_exploration_10042024_full$adjustedweight  / 2546) * 0.1818,
                                  ifelse(DS_individuals_exploration_10042024_full$Year >= 2015 & DS_individuals_exploration_10042024_full$Year <= 2016,
                                         (DS_individuals_exploration_10042024_full$adjustedweight  / 2723) * 0.1818,
                                         ifelse(DS_individuals_exploration_10042024_full$Year >= 2017 & DS_individuals_exploration_10042024_full$Year <= 2019,
                                                (DS_individuals_exploration_10042024_full$adjustedweight  / 3558) * 0.2727,
                                                NA))))  # NA for years outside specified ranges


mean_final_weight_per_year <- aggregate(FinalWeight ~ Year, DS_individuals_exploration_10042024_full, mean, na.rm = TRUE)
print(mean_final_weight_per_year)

```

## Differences-in-Differences Full dataset - Weights


```{r DiDestw, echo=FALSE}

library(stargazer)
library(tidyverse)
library(lmtest)

# using `srvyr`
library(srvyr)
library(survey)
DS_individuals_exploration_10042024_weighted <- DS_individuals_exploration_10042024_full %>% as_survey_design(ids = PSU, weights = FinalWeight)


did_model1_w <- svyglm(`Sum of Energykcal` ~ treatment * post_policy, family = gaussian(), design = DS_individuals_exploration_10042024_weighted)

did_model2_w <- svyglm(`Sum of EnergykJ` ~ treatment * post_policy, family = gaussian(), design = DS_individuals_exploration_10042024_weighted)

did_model3_w <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy, family = gaussian(), design = DS_individuals_exploration_10042024_weighted)

did_model4_w <- svyglm(`Sum of FoodEkJ` ~ treatment * post_policy, family = gaussian(), design = DS_individuals_exploration_10042024_weighted)


# Create a summary table
stargazer(did_model1_w, did_model2_w, did_model3_w, did_model4_w, type = "text", 
          title = "Difference-in-Differences Models",
          header = FALSE, 
          model.names = FALSE,
          column.labels = c("Energykcal", "EnergykJ", "FoodEkcal", "FoodEkJ"),
          covariate.labels = c("Treatment group", "Post-treatment", "Diff in Diff", "Intercept"),
          omit.stat = c("f", "ser"), # Omit statistics like F-statistic and standard error of regression
                    digits = 3) # Number of digits to display

```

## Differences-in-Differences Full dataset for variable FoodEkcal - Weights + Covariates

The lower AUC/BIC is model 1 

```{r DiDestw_c, echo=FALSE}

library(stargazer)
library(tidyverse)
library(lmtest)


did_model1_w_c <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy + Sex, family = gaussian(), design = DS_individuals_exploration_10042024_weighted)

did_model2_w_c <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy + Sex + AgeR, family = gaussian(), design = DS_individuals_exploration_10042024_weighted)

did_model3_w_c <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy + Sex + AgeR + region, family = gaussian(), design = DS_individuals_exploration_10042024_weighted)

did_model4_w_c <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy + Sex + AgeR + region + country, family = gaussian(), design = DS_individuals_exploration_10042024_weighted)

# Create a summary table

stargazer(did_model1_w_c, did_model2_w_c,did_model3_w_c, did_model4_w_c, type = "text", 
          title = "Difference-in-Differences Models",
          header = FALSE, 
          model.names = FALSE,
          column.labels = c("Control 1", "Control 2", "Control 3", "Control 4"),
          covariate.labels = c("Treatment", "Post policy","Sex", "Age", "Region","Country", "Dif. in changes over time", "Intercept - Baseline Average"),
          omit.stat = c("f", "ser"), # Omit statistics like F-statistic and standard error of regression
                    digits = 3) # Number of digits to display


#  Lower AIC/BIC
cat("AIC for Model 1:", AIC(did_model1_w_c), "\n")
cat("AIC for Model 2:", AIC(did_model2_w_c), "\n")
cat("AIC for Model 3:", AIC(did_model3_w_c), "\n")
cat("AIC for Model 4:", AIC(did_model4_w_c), "\n")

```

## Differences-in-Differences 18+ only - Weights


```{r DiDestadult_w, echo=FALSE}

#split the database

DS_individuals_exploration_10042024_full$AgeR <- as.numeric(as.character(DS_individuals_exploration_10042024_full$AgeR))

library(dplyr)
DS_individuals_exploration_10042024_filtered <- DS_individuals_exploration_10042024_full %>%
  dplyr::filter(AgeR >= 17)


# using `srvyr`
library(srvyr)
library(survey)
DS_individuals_exploration_10042024_weighted_16 <- DS_individuals_exploration_10042024_filtered %>% as_survey_design(ids = PSU, weights = FinalWeight)


#create a contingency table
cont_tablead <- table(DS_individuals_exploration_10042024_filtered$Year, DS_individuals_exploration_10042024_filtered$treatment)
cont_tablead
cont_table_bmiad <- table(DS_individuals_exploration_10042024_filtered$Year, DS_individuals_exploration_10042024_filtered$Income)
cont_table_bmiad
cont_table_polad <- table(DS_individuals_exploration_10042024_filtered$Year, DS_individuals_exploration_10042024_filtered$post_policy)
cont_table_polad

library(tidyverse)
library(lmtest)


did_model5_w <- svyglm(`Sum of Energykcal` ~ treatment * post_policy, family = gaussian(), design = DS_individuals_exploration_10042024_weighted_16)

did_model6_w <- svyglm(`Sum of EnergykJ` ~ treatment * post_policy, family = gaussian(), design = DS_individuals_exploration_10042024_weighted_16)

did_model7_w <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy, family = gaussian(), design = DS_individuals_exploration_10042024_weighted_16)

did_model8_w <- svyglm(`Sum of FoodEkJ` ~ treatment * post_policy, family = gaussian(), design = DS_individuals_exploration_10042024_weighted_16)

# Create a summary table
stargazer(did_model5_w, did_model6_w, did_model7_w, did_model8_w, type = "text", 
          title = "Difference-in-Differences Models, adults only",
          header = FALSE, 
          model.names = FALSE,
          column.labels = c("Energy Kl DiD", "Energy Kj DiD", "Food kl DiD", "Food kj DiD"),
          covariate.labels = c("Treatment group", "Post-treatment", "Diff in Diff", "Intercept"),
          omit.stat = c("f", "ser"), # Omit statistics like F-statistic and standard error of regression
          digits = 3) # Number of digits to display

```

## Differences-in-Differences 18+ only for variable FoodEkcal - Weights + Covariates

The lower AUC/BIC is model 1. 

```{r DiDestw_c_16, echo=FALSE}

library(stargazer)
library(tidyverse)
library(lmtest)

did_model5_w_c <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy + Sex, family = gaussian(), design = DS_individuals_exploration_10042024_weighted_16)

did_model6_w_c <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy + Sex + AgeR, family = gaussian(), design = DS_individuals_exploration_10042024_weighted_16)

did_model7_w_c <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy + Sex + AgeR + region, family = gaussian(), design = DS_individuals_exploration_10042024_weighted_16)

did_model8_w_c <- svyglm(`Sum of FoodEkcal` ~ treatment * post_policy + Sex + AgeR + region + country, family = gaussian(), design = DS_individuals_exploration_10042024_weighted_16)

# Create a summary table

stargazer(did_model6_w_c,did_model7_w_c, did_model8_w_c, did_model9_w_c, type = "text", 
          title = "Difference-in-Differences Models 16+ Only",
          header = FALSE, 
          model.names = FALSE,
          column.labels = c("Control 1", "Control 2", "Control 3", "Control 4"),
          covariate.labels = c("Treatment", "Post policy","Sex", "Age", "Region","Country", "Dif. in changes over time", "Intercept - Baseline Average"),
          omit.stat = c("f", "ser"), # Omit statistics like F-statistic and standard error of regression
                    digits = 3) # Number of digits to display

#  Lower AIC/BIC
cat("AIC for Model 1:", AIC(did_model5_w_c), "\n")
cat("AIC for Model 2:", AIC(did_model6_w_c), "\n")
cat("AIC for Model 3:", AIC(did_model7_w_c), "\n")
cat("AIC for Model 4:", AIC(did_model8_w_c), "\n")

```

## Check for Parallel Trends Assumption - Weights


```{r DiDplotw, echo=FALSE}

# Adjust the outcome by the mean of FinalWeight for each group and year
adjusted_plot_data <- DS_individuals_exploration_10042024_weighted_16 %>%
  group_by(Year, treatment) %>%
  summarise(mean_outcome = mean(`Sum of Energykcal`, na.rm = TRUE),
            mean_final_weight = mean(FinalWeight, na.rm = TRUE)) %>%
  mutate(adjusted_mean_outcome = mean_outcome / mean_final_weight) %>%
  ungroup()

# Plot the adjusted mean outcomes
ggplot(adjusted_plot_data, aes(x = Year, y = adjusted_mean_outcome, color = as.factor(treatment))) +
  geom_line() +
  labs(title = "Pre-Treatment Trends Adjusted by Final Weight",
       y = "Adjusted Mean of Energykcal",
       x = "Year",
       color = "Treatment Group")

# Adjust the outcome by the mean of FinalWeight for each group and year
adjusted_plot_data2 <- DS_individuals_exploration_10042024_weighted_16 %>%
  group_by(Year, treatment) %>%
  summarise(mean_outcome = mean(`Sum of EnergykJ`, na.rm = TRUE),
            mean_final_weight = mean(FinalWeight, na.rm = TRUE)) %>%
  mutate(adjusted_mean_outcome = mean_outcome / mean_final_weight) %>%
  ungroup()

# Plot the adjusted mean outcomes
ggplot(adjusted_plot_data2, aes(x = Year, y = adjusted_mean_outcome, color = as.factor(treatment))) +
  geom_line() +
  labs(title = "Pre-Treatment Trends Adjusted by Final Weight",
       y = "Adjusted Mean of Energykj",
       x = "Year",
       color = "Treatment Group")

# Adjust the outcome by the mean of FinalWeight for each group and year
adjusted_plot_data3 <- DS_individuals_exploration_10042024_weighted_16 %>%
  group_by(Year, treatment) %>%
  summarise(mean_outcome = mean(`Sum of FoodEkcal`, na.rm = TRUE),
            mean_final_weight = mean(FinalWeight, na.rm = TRUE)) %>%
  mutate(adjusted_mean_outcome = mean_outcome / mean_final_weight) %>%
  ungroup()

# Plot the adjusted mean outcomes
ggplot(adjusted_plot_data3, aes(x = Year, y = adjusted_mean_outcome, color = as.factor(treatment))) +
  geom_line() +
  labs(title = "Pre-Treatment Trends Adjusted by Final Weight",
       y = "Adjusted Mean of FoodEkcal",
       x = "Year",
       color = "Treatment Group")

# Adjust the outcome by the mean of FinalWeight for each group and year
adjusted_plot_data4 <- DS_individuals_exploration_10042024_weighted_16 %>%
  group_by(Year, treatment) %>%
  summarise(mean_outcome = mean(`Sum of FoodEkJ`, na.rm = TRUE),
            mean_final_weight = mean(FinalWeight, na.rm = TRUE)) %>%
  mutate(adjusted_mean_outcome = mean_outcome / mean_final_weight) %>%
  ungroup()

# Plot the adjusted mean outcomes
ggplot(adjusted_plot_data4, aes(x = Year, y = adjusted_mean_outcome, color = as.factor(treatment))) +
  geom_line() +
  labs(title = "Pre-Treatment Trends Adjusted by Final Weight",
       y = "Adjusted Mean of FoodEkJ",
       x = "Year",
       color = "Treatment Group")




```

