---
title: "National Diet and Nutrition Survey_Two-way Fixed Effects"
author: "Constanza Avalos"
date: "2024-04-17"
output: 
  html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

## Datasets 2008-2019

This survey contains various datasets, with the food diary report being the focus of my analysis. The food diary report is a self-reported record of all items consumed over a 24-hour period (midnight to midnight) on the previous day. Participants were required to complete four recalls on different days that were not consecutive. There are three primary datasets: one categorizing consumption reports into food categories, another detailing grams consumed per category at an individual level, and finally one indicating nutrient intake at the individual level. These databases include information such as reporting days of the week, time, date, country, age, gender and others. Moreover, they have a multilevel structure encompassing individuals,,households, regions and countries.The data also encompasses socio-demographic details along with anthropometric measurements (BMI, weight, height) at an individual level. Models incorporate survey weights.

## Treatment groups

The dataset "ndns_rp_yr9_yr9_11a_dayleveldietarydata_nutrients_uk_20210831" contains information on nutrient intake at the individual level. I plan to divide the data for analysis, specifically focusing on individuals aged 18 and older. Additionally, this database includes the variable region, which is used to differentiate between the treatment and control groups.

In this quasi-experimental design, individuals from Northern Ireland (D=1) are classified as part of the treatment group while others are assigned to the control group (D=0).

According to existing literature, there were variations in how quickly this policy was implemented by retailers since 2013. Considering that Ireland is geographically distant from mainland United Kingdom, we infer a delay in implementing the labeling policy. Ireland is region number 6 in the dataset.

Individuals in the lower income_tertile (D=1) are included in the treatment group, while others are placed in the control group (D=0) as part of the second quasi-experimental design. Previous research suggests that lower socioeconomic groups have benefited most since retailers implemented this policy. Unlike taxes, food labeling does not place an economic burden on consumers but instead guides them toward healthy eating behavior.


```{r data, echo=FALSE}

# Working directory

setwd("C:/Users/n51867ca/OneDrive - The University of Manchester/R/Second paper")

# input Stata file
library(openxlsx)
library(readxl)
library(foreign)
library(stringi)
library(readr)

#library(readr)
DS_individuals_exploration_10042024_SCM_region <- read_csv("DS_individuals_exploration_10042024_SCM_region.csv")
DS_individuals_exploration_10042024_SCM_SES <- read_csv("DS_individuals_exploration_10042024_SCM_SES.csv")


# Data structure

head(DS_individuals_exploration_10042024_SCM_region)
head(DS_individuals_exploration_10042024_SCM_SES)


```
## Statistics 

```{r summary, echo=FALSE, results='asis'}

library(summarytools)

library(tidyverse)
library(lmtest)
library(dplyr)

DS_individuals_exploration_10042024_SCM_region <- DS_individuals_exploration_10042024_SCM_region %>%
    mutate(Year = as.numeric(Year), # Convert SurveyYear to numeric
          Year = case_when(
           Year == 1 ~ 2009,
           Year == 2 ~ 2010,
           Year == 3 ~ 2011,
           Year == 4 ~ 2012,
           Year == 5 ~ 2013,
           Year == 6 ~ 2014,
           Year == 7 ~ 2015,
           Year == 8 ~ 2016,
           Year == 9 ~ 2017,
           Year == 10 ~ 2018,
           Year == 11 ~ 2019,
    TRUE ~ Year # Keeps the original value if none of the conditions above are met
  ))

DS_individuals_exploration_10042024_SCM_SES <- DS_individuals_exploration_10042024_SCM_SES %>%
    mutate(Year = as.numeric(Year), # Convert SurveyYear to numeric
          Year = case_when(
           Year == 1 ~ 2009,
           Year == 2 ~ 2010,
           Year == 3 ~ 2011,
           Year == 4 ~ 2012,
           Year == 5 ~ 2013,
           Year == 6 ~ 2014,
           Year == 7 ~ 2015,
           Year == 8 ~ 2016,
           Year == 9 ~ 2017,
           Year == 10 ~ 2018,
           Year == 11 ~ 2019,
    TRUE ~ Year # Keeps the original value if none of the conditions above are met
  ))


# Replace negative values with NA across the entire dataset
DS_individuals_exploration_10042024_SCM_region[DS_individuals_exploration_10042024_SCM_region < 0] <- NA
DS_individuals_exploration_10042024_SCM_SES[DS_individuals_exploration_10042024_SCM_SES < 0] <- NA

summary <- dfSummary(DS_individuals_exploration_10042024_SCM_region, graph.magnif = 0.75, plain.ascii = FALSE, style = "grid")
print(summary, method = 'render')

summary2 <- dfSummary(DS_individuals_exploration_10042024_SCM_SES, graph.magnif = 0.75, plain.ascii = FALSE, style = "grid")
print(summary2, method = 'render')

```


## Two-way Fixed Effects - Region


```{r 2fixed, echo=FALSE}

# Install and load the plm package
library(plm)
library(stargazer)

# Convert data to a pdata.frame for panel data analysis
data_2fixed <- pdata.frame(DS_individuals_exploration_10042024_SCM_region, index = c("Region", "Year"))

# Define the model
model_2fixed <- plm(Sum.of.Energykcal ~ Average.of.bmival2 + Average.of.AgeR + Average.of.Income, data = data_2fixed, model = "within", effect = "twoways")
model_2fixed2 <- plm(Sum.of.EnergykJ ~ Average.of.bmival2 + Average.of.AgeR + Average.of.Income, data = data_2fixed, model = "within", effect = "twoways")
model_2fixed3 <- plm(Sum.of.FoodEkcal ~ Average.of.bmival2 + Average.of.AgeR + Average.of.Income, data = data_2fixed, model = "within", effect = "twoways")
model_2fixed4 <- plm(Sum.of.FoodEkJ ~ Average.of.bmival2 + Average.of.AgeR + Average.of.Income, data = data_2fixed, model = "within", effect = "twoways")

# Summary of the model

# Create a summary table
stargazer(model_2fixed, model_2fixed2, model_2fixed3, model_2fixed4, type = "text", 
          title = "Two-way Fixed Effects Models - Region",
          header = FALSE, 
          model.names = FALSE,
          column.labels = c("Energykcal", "EnergykJ", "FoodEkcal", "FoodEkJ"),
          covariate.labels = c("Average.of.bmival2", "Average.of.AgeR"),
          omit.stat = c("f", "ser"), # Omit statistics like F-statistic and standard error of regression
                    digits = 3) # Number of digits to display

```

## Plotting Region


```{r plot, echo=FALSE}

# Load necessary libraries
library(plm)
library(dplyr)
library(ggplot2)


# Extract fitted values directly aligned with pdata
data_2fixed$fitted_values <- fitted(model_2fixed)

# Create a data frame for plotting that includes Year, Region, and Fitted Values
plot_data <- data_2fixed %>% 
  as.data.frame() %>%
  select(Year, Region, fitted_values)

# Plotting
ggplot(plot_data, aes(x = Year, y = fitted_values, group = Region, color = Region)) +
  geom_line() +
  geom_point() +
  labs(title = "Predicted Weighted Sum of Energykcal Across Years",
       x = "Year",
       y = "Predicted Weighted Sum of Energykcal") +
  theme_minimal()


# Extract fitted values directly aligned with pdata
data_2fixed$fitted_values <- fitted(model_2fixed2)

# Create a data frame for plotting that includes Year, Region, and Fitted Values
plot_data2 <- data_2fixed %>% 
  as.data.frame() %>%
  select(Year, Region, fitted_values)

# Plotting
ggplot(plot_data2, aes(x = Year, y = fitted_values, group = Region, color = Region)) +
  geom_line() +
  geom_point() +
  labs(title = "Predicted Weighted Sum of Energykj Across Years",
       x = "Year",
       y = "Predicted Weighted Sum of Energykcal") +
  theme_minimal()

# Extract fitted values directly aligned with pdata
data_2fixed$fitted_values <- fitted(model_2fixed3)

# Create a data frame for plotting that includes Year, Region, and Fitted Values
plot_data3 <- data_2fixed %>% 
  as.data.frame() %>%
  select(Year, Region, fitted_values)

# Plotting
ggplot(plot_data3, aes(x = Year, y = fitted_values, group = Region, color = Region)) +
  geom_line() +
  geom_point() +
  labs(title = "Predicted Weighted Sum of Foodkcal Across Years",
       x = "Year",
       y = "Predicted Weighted Sum of Energykcal") +
  theme_minimal()

# Extract fitted values directly aligned with pdata
data_2fixed$fitted_values <- fitted(model_2fixed4)

# Create a data frame for plotting that includes Year, Region, and Fitted Values
plot_data4 <- data_2fixed %>% 
  as.data.frame() %>%
  select(Year, Region, fitted_values)

# Plotting
ggplot(plot_data4, aes(x = Year, y = fitted_values, group = Region, color = Region)) +
  geom_line() +
  geom_point() +
  labs(title = "Predicted Weighted Sum of Foodkcal Across Years",
       x = "Year",
       y = "Predicted Weighted Sum of Energykcal") +
  theme_minimal()




```

## Two-way Fixed Effects - Income


```{r 2fixedses, echo=FALSE}

# Install and load the plm package
library(plm)
library(stargazer)

# Convert data to a pdata.frame for panel data analysis
data_2fixed_ses <- pdata.frame(DS_individuals_exploration_10042024_SCM_SES, index = c("Income", "Year"))

# Define the model
model_2fixed_ses <- plm(Average.of.Sum.of.Energykcal2 ~ Average.of.bmival22 + Average.of.AgeR, data = data_2fixed_ses, model = "within", effect = "twoways")
model_2fixed2_ses <- plm(Average.of.Sum.of.EnergykJ2 ~ Average.of.bmival22 + Average.of.AgeR, data = data_2fixed_ses, model = "within", effect = "twoways")
model_2fixed3_ses <- plm(Average.of.Sum.of.FoodEkcal2 ~ Average.of.bmival22 + Average.of.AgeR, data = data_2fixed_ses, model = "within", effect = "twoways")
model_2fixed4_ses <- plm(Average.of.Sum.of.FoodEkJ2 ~ Average.of.bmival22 + Average.of.AgeR, data = data_2fixed_ses, model = "within", effect = "twoways")

# Summary of the model

# Create a summary table
stargazer(model_2fixed_ses, model_2fixed2_ses, model_2fixed3_ses, type = "Text", 
          title = "Two-way Fixed Effects Models - Income",
          header = FALSE, 
          model.names = FALSE,
          column.labels = c("Energykcal", "EnergykJ", "FoodEkcal"),
          covariate.labels = c("Average.of.bmival2", "Average.of.AgeR"),
          omit.stat = c("f", "ser"), # Omit statistics like F-statistic and standard error of regression
                    digits = 4) # Number of digits to display


```
## Plotting Income


```{r plotses, echo=FALSE}

# Load necessary libraries
library(plm)
library(dplyr)
library(ggplot2)


# Extract fitted values directly aligned with pdata
pdf("path_plot2.pdf", width = 7, height = 5)
data_2fixed_ses$fitted_values <- fitted(model_2fixed_ses)

# Create a data frame for plotting that includes Year, Region, and Fitted Values
plot_data5 <- data_2fixed_ses %>% 
  as.data.frame() %>%
  select(Year, Income, fitted_values)

# Plotting
ggplot(plot_data5, aes(x = Year, y = fitted_values, group = Income, color = Income)) +
  geom_line() +
  geom_point() +
  labs(title = "Predicted Weighted Sum of Energykcal Across Years",
       x = "Year",
       y = "Predicted Weighted Sum of Energykcal") +
  theme_minimal()
dev.off()

pdf("path_plot3.pdf", width = 7, height = 5)
# Extract fitted values directly aligned with pdata
data_2fixed_ses$fitted_values <- fitted(model_2fixed2_ses)

# Create a data frame for plotting that includes Year, Region, and Fitted Values
plot_data6 <- data_2fixed_ses %>% 
  as.data.frame() %>%
  select(Year, Income, fitted_values)

# Plotting
ggplot(plot_data6, aes(x = Year, y = fitted_values, group = Income, color = Income)) +
  geom_line() +
  geom_point() +
  labs(title = "Predicted Weighted Sum of Energykj Across Years",
       x = "Year",
       y = "Predicted Weighted Sum of Energykj") +
  theme_minimal()

dev.off()

# Extract fitted values directly aligned with pdata
data_2fixed_ses$fitted_values <- fitted(model_2fixed3_ses)

# Create a data frame for plotting that includes Year, Region, and Fitted Values
plot_data7 <- data_2fixed_ses %>% 
  as.data.frame() %>%
  select(Year, Income, fitted_values)

# Plotting
ggplot(plot_data7, aes(x = Year, y = fitted_values, group = Income, color = Income)) +
  geom_line() +
  geom_point() +
  labs(title = "Predicted Weighted Sum of FoodEkcal2 Across Years",
       x = "Year",
       y = "Predicted Weighted Sum of Energykcal") +
  theme_minimal()





```
